---
title: "inferential statistics assignment 1"
author: "Bowen Liu"
date: "February 19, 2016"
output: pdf_document
---
# Investigate the exponential distribution in R and compare it with the Central Limit Theorem.
The exponential distribution: rexp(n, lambda)
The rate parameter of lamba = 0.2
The mean of exponential distribution is 1/lambda = 1/0.2 = 5
The standard deviation is also 1/lambda = 5
The sample size is 40
The simulation size is 1000

```{r}
set.seed(1000) # set the seed for replicating the experiment
simu_size <- 1000
sample_size <- 40
lambda <- 0.2
# simulate 1000 times of sampling
simu_data <- replicate(simu_size, rexp(sample_size, lambda))
# create a dataframe to store means and variance of sampling and population
df <- data.frame(Mean=c(mean(colMeans(simu_data)), 1/lambda),
                 Variance=c(var(colMeans(simu_data)), 1/lambda^2))
rownames(df) <- c("sample", "polulation")
df
```

as we can see, the mean of sampling distribution of sample mean is 4.986963, pretty close to
the population mean of 5. And the the variance of sampling distribution of sample
mean is 0.654343 is, very close to the population variance of 25/sqrt(40)

```{r,echo=FALSE}
library(ggplot2)
sample_means <- colMeans(simu_data)
mean_of_sample_means <- mean(sample_means)
sd_of_sample_means <- sd(sample_means)
zscore <- (sample_means - mean_of_sample_means) / sd_of_sample_means
qplot(zscore, geom = "blank") +
    geom_line(aes(y = ..density.., colour = 'Sample Distribution'), stat = 'density') +  
    stat_function(fun = dnorm, aes(colour = 'Normal Distribution')) +
    scale_colour_manual(name = 'Comparision', values = c('red', 'blue')) +
    ylab("density") + xlab("z-score") + ggtitle("Sampling Distribution of Sample Means") 
```

we can see from the plotting, the sample distribution is very close to the normal distribution.