---
title: "inferential statistics assignment 2"
author: "Bowen Liu"
date: "February 20, 2016"
output: pdf_document
---

# Analyze the ToothGrowth data in the R datasets package.

## 1. Load the ToothGrowth data and perform some basic exploratory data analyses

```{r}
library(ggplot2)
data("ToothGrowth")
str(ToothGrowth)
ggplot(ToothGrowth, aes(dose, len)) +
    geom_point(aes(shape = supp))
```

It appears that there is a positive correlation between dose and tooth length

## 2. Provide a basic summary of the data.

```{r}
summary(ToothGrowth)
summary(lm(ToothGrowth$len ~ ToothGrowth$dose))
```

## 3. Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose.
```{r}
# null hypothesis 1: supplements have the same tooth growth effect
hypoth1<-t.test(len ~ supp, data = ToothGrowth)
hypoth1$conf.int
hypoth1$p.value
```

In hypothesis 1, the confidence interval covers 0 and the pvalue is greater than 0.05. We cannot reject null hypothesis1


In hypothesis 2, we set null hypothesis with different dose to if the supplements have the same tooth growth effect

```{r}
df <- data.frame()
lvs <- levels(factor(ToothGrowth$dose))
for (lv in lvs) {
    hypoth2 <- t.test(len ~ supp, data = subset(ToothGrowth, dose == lv))
    df <- rbind(df, c(hypoth2$conf.int, hypoth2$p.value))
}
colnames(df) <- c("-2sigma", "2sigma", "p-value")
rownames(df) <- c("dose=0.5", "dose=1", "dose=2")
df
```

In hypothesis 2, only dose=2 covering 0 value and with p-value > 0.05, accepts the null hypothese.

## 4. State your conclusions and the assumptions needed for your conclusions.
Orange juice and ascorbic acid could have different tooth growth for doses of 0.5 and 1.0. And they could have the same tooth growth for dose of 2.0.

Assumptions:

-Tooth length conforms to T distribution

-No other factors affect tooth length